Notes on conformal manifolds

### 2025-09-05 Draft_MLFlatDirections.tex
Modifications during meeting.

The AdS/CFT correspondence stands as one of the most profound dualities in theoretical physics, establishing a remarkable equivalence between string theory solutions on anti-de Sitter (AdS) spacetimes and conformal field theories (CFTs) living on their boundaries~\cite{Maldacena:1997re}. This correspondence has revolutionized our understanding of both quantum gravity and strongly coupled field theories, providing unprecedented insights into the holographic nature of gravity. Within this holographic framework, supergravity theories in AdS backgrounds serve as the low-energy effective descriptions of string theory compactifications, making them natural laboratories for exploring the gravitational side of the duality. The CFT side plays a crucial role in statistical physics, as CFT are known to describe phase transitions ; they are also fixed point of Renormalisation Group flow and this is this latter aspect that is our interest in this paper. An interesting question is indeed to know whether those CFT are isolated fixed points of RG flows, or belong to a continuous family of CFT. If this is the case, the space of continuous deformations that takes from one CFT to another is called a conformal manifold. These deformations are parametrised by exactly marginal operators, or in other words operators that preserves the conformal symmetry, \textit{i.e.} whose $\beta$ functions exactly vanish.

@@

In this work, we demonstrate the viability of the aforementioned machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of six-dimensional non-chiral $\mathcal{N} = (1,1)$ supergravity on $AdS_3 \times S^3$, or to type IIB supergravity on $AdS_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the conformal manifold with a symbolic regression technique. There exists a large body of literature on symbolic regression, using methods from genetic programming \cite{koza1994genetic,virgolin2021improving,randall2022bingo,burlacu2019parsimony}, to deep learning \cite{petersen2019deep,kamienny2022end}, generative models \cite{valipour2021symbolicgpt}, diffusion models \cite{bastiani2025diffusion}, and equation learning with the nodes being symbolic operations \cite{2018arXiv180607259S}. An state-of-the-art algorithm is AIFeynman \cite{Udrescu:2019mnk}. It uses neural networks to identify structures in the dataset (such as translational symmetries, multiplicative separability, compositionality\dots) to recursively define simpler problems on which they can fit the solutions with polynomials.  However, this method is slow, usually limited to low-dimensional spaces, and, by construction, can only fit one polynomial at a time. Alternatively, we develop here an algorithm based on symbolic regression using Annealing Importance Sampling (AIS) \cite{neal1998annealedimportancesampling}, since this approach is much better suited to the study of geometric loci defined by intersecting polynomials in high-dimensional spaces

@@

In this work, we aim to identify such solutions. 
A preliminary numerical exploration suggests that attention can be focused on the five scalars $x_{1}, x_{2}, x_{4}, \tilde{x}_{8}$, and $x_{10}$. 
The solutions we obtain in this truncation remain valid solutions of the complete theory, since we first compute $\nabla V$ with all scalar fields included, \textit{i.e.} using eq.~\eqref{eq:scalarpotential}, and only then setting the remaining fields,
\begin{equation} \label{eq:defvecy}
	\vec{y} = (x_3,x_5,x_6,\tilde{x}_7,\tilde{x}_9,x_{11},x_{12},\tilde{x}_{13}),
\end{equation}
to zero, as detailed in sec.~\ref{sec:graddes}. By performing the differentiation before the truncation, we ensure that the resulting configurations satisfy the full equations of motion and are therefore legitimate solutions of the complete theory.

@@

Let us now explain in more detail how does this procedure goes. The goal is to reconstruct some distribution function $d(z)$, where here $z$ in going to be some polynomial. We will try to reconstruct this density function by series of density function $\pi_n(z_n) = \gamma_n(z_n)/Z_n$ with $n = 1,\dots,N$ the number of annealing steps, $\pi_n$ is defined in terms of an unnormalized density $\gamma_n$ and we have the normalizing constant $Z_n = \int \gamma_n(z) \mathrm{d}z$. We also assume we have a sequence of inverse temperature constants $\beta_n$, where $0 = \beta_1 < \beta_2 <\dots< \beta_N = 1$. The unnormalized density at level $n$ is then given in terms of a prior distribution $p_0(z)$ over the hypothesis space and a loss function $L(z)$:
		\begin{equation}\label{eq:gamman}
			\gamma_n(z) := d_0(z) \: \exp \left( -\beta_n L(z) \right).
		\end{equation}

>>>>

The AdS/CFT correspondence stands as one of the most profound dualities in theoretical physics, establishing a remarkable equivalence between string theory solutions on anti-de Sitter (AdS) spacetimes and conformal field theories (CFTs) living on their boundaries~\cite{Maldacena:1997re}. This correspondence has revolutionized our understanding of both quantum gravity and strongly coupled field theories, providing unprecedented insights into the holographic nature of gravity. Within this holographic framework, supergravity theories in AdS backgrounds serve as the low-energy effective descriptions of string theory compactifications, making them natural laboratories for exploring the gravitational side of the duality. The CFT side plays a crucial role in statistical physics, as CFT are known to describe phase transitions ; they are also fixed point of Renormalisation Group flow and this is this latter aspect that is our interest in this paper. An interesting question is indeed to know whether those CFT are isolated fixed points of RG flows, or belong to a continuous family of CFT. If this is the case, the space of continuous deformations that takes from one CFT to another is called a conformal manifold. These deformations are parametrised by exactly marginal operators, or in other words operators that preserves the conformal symmetry, \textit{i.e.} whose $\beta$ functions exactly vanish. \ce{Needs some modifications.}

@@

In this work, we demonstrate the viability of the aforementioned machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of six-dimensional non-chiral $\mathcal{N} = (1,1)$ supergravity on $AdS_3 \times S^3$, or to type IIB supergravity on $AdS_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the conformal manifold with a symbolic regression technique. There exists a large body of literature on symbolic regression, using methods from genetic programming \cite{koza1994genetic,virgolin2021improving,randall2022bingo,burlacu2019parsimony}, to deep learning \cite{petersen2019deep,kamienny2022end}, generative models \cite{valipour2021symbolicgpt}, diffusion models \cite{bastiani2025diffusion}, and equation learning with the nodes being symbolic operations \cite{2018arXiv180607259S}. An state-of-the-art algorithm is AIFeynman \cite{Udrescu:2019mnk}. It uses neural networks to identify structures in the dataset (such as translational symmetries, multiplicative separability, compositionality\dots) to recursively define simpler problems on which they can fit the solutions with polynomials.  However, this method is slow, usually limited to low-dimensional spaces, and, by construction, can only fit one polynomial at a time. Alternatively, we develop here a symbolic regression algorithm based on Annealing Importance Sampling (AIS) \cite{neal1998annealedimportancesampling}, since this approach is much better suited to the study of geometric loci defined by intersecting polynomials in high-dimensional spaces

@@

In this work, we aim to identify such solutions. 
A preliminary numerical exploration suggests that attention can be focused on the five scalars $x_{1}, x_{2}, x_{4}, \tilde{x}_{8}$, and $x_{10}$. 
To ensure that the solutions we obtain in this truncation remain valid solutions of the complete theory, we first compute $\nabla V$ with all scalar fields included, \textit{i.e.} using eq.~\eqref{eq:scalarpotential}, and only then setting the remaining fields,
\begin{equation} \label{eq:defvecy}
	\vec{y} = (x_3,x_5,x_6,\tilde{x}_7,\tilde{x}_9,x_{11},x_{12},\tilde{x}_{13}),
\end{equation}
to zero, as detailed in sec.~\ref{sec:graddes}. By performing the differentiation before the truncation, we ensure that the resulting configurations satisfy the full equations of motion and are therefore legitimate solutions of the complete theory.

@@

Let us now explain in more detail how does this procedure goes. The goal is to reconstruct some distribution function $d(z)$, where here $z$ in going to be some polynomial. We will try to reconstruct this density function by series of density function $\pi_n(z_n) = \gamma_n(z_n)/Z_n$ with $n = 1,\dots,N$ the number of annealing steps, $\pi_n$ is defined in terms of an unnormalized density $\gamma_n$ and we have the normalizing constant $Z_n = \int \gamma_n(z) \mathrm{d}z$. We also assume we have a sequence of inverse temperature constants $\beta_n$, where $0 = \beta_1 < \beta_2 <\dots< \beta_N$. The unnormalized density at level $n$ is then given in terms of a prior distribution $p_0(z)$ over the hypothesis space and a loss function $L(z)$:
		\begin{equation}\label{eq:gamman}
			\gamma_n(z) := d_0(z) \: \exp \left( -\beta_n L(z) \right).
		\end{equation}
